---
title: "Numpy 用法"
author: "Jin"
date: "2020年6月"
institute: 中南财经政法大学统计与数学学院
csl: ./style/chinese-gb7714-2015-numeric.csl
css: ./style/markdown.css
bibliography: [./Bibfile.bib]
eqnPrefixTemplate: ($$i$$)
link-citations: true
linkReferences: true
chapters: true
tableEqns: false
autoEqnLabels: false
classoption: "aspectratio=1610"
---

```{r setup, echo=F, purl=F}
knitr::opts_knit$set(root.dir = getwd())
knitr::opts_chunk$set(echo = TRUE, results = 'hide')
knitr::opts_chunk$set(warning = FALSE, message=FALSE)
knitr::opts_chunk$set(fig.align="center"
                      ## ,out.width="0.9\\textwidth" # latex
                      ,out.width="80%" # for both latex and html
                      ,fig.width=5, fig.height=3
                      )
```

```{r prepare, echo=F, purl=F}
rm(list=ls())
options(digits=4)
options(scipen=100)
graphics.off()
Sys.setlocale("LC_ALL", "Chinese")
library(reticulate)
```


# Introduction


### Facts

1.  Initial release: As Numeric, 1995; as NumPy, 2006
2.  Stable release: 1.11.2 / 3 October 2016;
3.  Website: <http://www.numpy.org>
4.  History: <https://en.wikipedia.org/wiki/NumPy>

### What is NumPy?

1.  NumPy is the fundamental package for scientific computing in Python.
2.  It is a Python library that provides a multidimensional array
    object, various derived objects (such as masked arrays and
    matrices), and an assortment of routines for fast operations on
    arrays, including


1.  mathematical,
2.  logical,
3.  shape manipulation,
4.  sorting, selecting, I/O,
5.  discrete Fourier transforms, basic linear algebra,
6.  basic statistical operations, random simulation and much more.

### NumPy

1.  At the core of the NumPy package, is the **ndarray** object.
2.  This encapsulates n-dimensional arrays of homogeneous data types,
    with many operations being performed in compiled code for
    performance.
3.  The points about sequence size and speed are particularly important
    in scientific computing.

### ndarray object

1.  Vectorization:Vectorization describes the absence of any explicit
    looping, indexing, etc., in the code - these things are taking
    place, of course, just "behind the scenes" in optimized,
    pre-compiled C code.
2.  Broadcasting:Broadcasting is the term used to describe the implicit
    element-by-element behavior of operations.
3.  NumPy fully supports an object-oriented approach with ndarray.
    ndarray is a class, possessing numerous methods and attributes.

# Basics

## Array Creation

### `array` function

1.  create an array from a regular Python list or tuple using the
    **`array`** function. The type of the resulting array is deduced
    from the type of the elements in the sequences.
2.  array transforms sequences of sequences into two-dimensional arrays,
    sequences of sequences of sequences into three-dimensional arrays,
    and so on.
3.  Often, the elements of an array are originally unknown, but its size
    is known.

### arrays with initial placeholder content

1.  The function **`zeros`** creates an array full of zeros,
2.  the function **`ones`** creates an array full of ones,
3.  the function **`empty`** creates an array whose initial content is
    random and depends on the state of the memory.
4.  the function `diag` creates the diagonal array,
5.  the function **`eye`** or **`identity`** creates an array with ones
    on the diagonal and zeros elsewhere.

### `arange` and `linspace` function

1.  **`arange`**:
2.  **`linspace`**:

### ndarray对象的属性

1.  ndarray.ndim
2.  ndarray.shape
3.  ndarray.size
4.  ndarray.dtype
5.  ndarray.itemsize
6.  ndarray.data

## 基本操作

### 基本操作

1.  数组上的算术运算符按元素方式应用。创建一个新的数组并填充结果。
2.  矩阵乘积可使用点函数或方法执行:
3.  许多一元运算，比如计算数组中所有元素的总和，都是作为ndarray类的方法实现的。
4.  通过指定轴参数，可以应用一个沿着数组的指定轴进行的操作(类似于R中的apply函数)。

### 例子

``` {python}
import numpy as np
a=np.arange(4)
b=np.array([2,5,8,9])
a*b

A=np.arange(12).reshape(3,4)
B=np.arange(13,25).reshape(4,3)
np.dot(A, B)
A.dot(B)
A.sum()
A.sum(axis=0)
A.sum(axis=1)

"""END"""
```

### 通用函数

1.  NumPy提供了熟悉的数学函数，如sin、cos和exp。
2.  在NumPy中，这些函数称为\“通用函数\”(ufunc)。
3.  在NumPy中，这些函数按元素对数组进行操作，生成一个数组作为输出。

``` {python}
A=np.arange(12).reshape(3,4)
np.exp(A)
np.sqrt(A)

"""END"""
```

## 索引、切片和迭代

### 索引、切片和迭代

1.  一维数组可以被索引、切片和迭代，就像列表和其他Python序列一样。

``` {python}
x=np.arange(12)**2
x[3]
x[2:6]
x[7:]
x[::-1]
x[9:2:-3]

"""END"""
```

### 索引、切片和迭代

1.  多维数组每个轴可以有一个索引。这些索引以元组的形式给出，中间用逗号分隔。
2.  当提供的索引少于轴的数量时，缺失的索引被认为是完整的片。

``` {python}
A=np.arange(24).reshape(4,6)
A[2,3]
A[1:3, 2:4]
A[1]
A[:, 2:4]
A[..., 3]

"""END"""
```

### 索引、切片和迭代

1.  点(…)表示生成完整索引元组所需的冒号。
2.  例如，如果x是一个有5个轴的数组，那么

-   `x[1, 2, ...]` is equivalent to `x[1, 2, :, :, :]`,
-   `x[..., 3]` to `x[:, : ,: ,:, 3]`
-   `x[4, ..., 5, :]` to `x[4, :, :, 5, :]`

### Indexing, Slicing and Iterating

1.  Iterating over multidimensional arrays is done with respect to
    the first axis
2.  if one wants to perform an operation on each element in the array,
    one can use the flat attribute which is an iterator over all the
    elements of the array

``` {python}
import numpy as np
A=np.arange(24).reshape(4,6)
for i in A:
    """打印A的各行"""
    print(i) 

for i in A.flat:
    """打印A中的每个元素"""
    print(i)

"""END"""
```

# Shape Manipulation

### Changing the shape of an array

1.  An array has a shape given by the number of elements along each axis
2.  The shape of an array can be changed with various commands.
3.  Note that the following three commands all return a modified array,
    but do not change the original array:

<!-- -->

1.  `ndarray.ravel(), ndarray.T, ndarry.reshape`
    1.  the `ndarray.resize` method modifies the array itself

``` {python}
import numpy as np
a = np.floor(10 * np.random.random((3,4)))
a.shape



a.ravel()
a.T
a.reshape(2,6)
a.resize(2,6)

"""END"""
```

### Stacking together different arrays

1.  `hstack, vstack`
2.  `column_stack, row_stack`
3.  `concatenate`
4.  `c_, r_`

### Splitting one array into several smaller ones

1.  `hsplit`
2.  `vsplit`
3.  `array_split`

# Fancy indexing and index tricks

### Indexing with Arrays of Indices-1D

1.  使用 `np.array` 对象作为索引下标可以取非连续元素

``` {python}
a = np.arange(12) ** 2 # the first 12 square numbers
i = np.array( [ 1,1,3,8,5 ] ) # an array of indices
a[i] # the elements of a at the positions i
np.array([ 1, 1, 9, 64, 25])

j = np.array( [ [ 3, 4], [ 9, 7 ] ] )
a[j]



"""END"""
```

### Indexing with Arrays of Indices-2D

1.  We can also give indexes for more than one dimension. The
    arrays of indices for each dimension must have the same shape.
2.  Naturally, we can put i and j in a sequence (say a list) and then do
    the indexing with the list.

``` {python}
a = np.arange(12).reshape(3,4)
i = np.array([[0,1], [1,2]])
j = np.array([[2,1], [3,3]])

a[i]
a[i,j]
a[i, 2]
a[:,j]

L = [i,j]
a[L]

"""END"""
```

### Indexing with Boolean Arrays

1.  use boolean arrays that have the same shape as the original array
2.  for each dimension of the array we give a 1D boolean array selecting
    the slices we want
3.  Note that the length of the 1D boolean array must coincide with the
    length of the dimension (or axis) you want to slice.

``` {python}
a = np.arange(12).reshape(3,4)
b=a>4
a[b]
a[b]=0

a = np.arange(12).reshape(3,4)
b1 = np.array([False,True,True])
b2 = np.array([True,False,True,False])
a[b1,:]
a[b1]
a[:,b2]
a[b1,b2]

"""END"""
```

### Indexing with strings

1.  Structured arrays are ndarrays whose datatype is a composition of
    simpler datatypes organized as a sequence of named fields.
2.  You can access and modify individual fields of a structured array by
    indexing with the field name.
3.  One can index and assign to a structured array with a multi-field
    index, where the index is a list of field names.

``` {python}
x = np.array([('Rex', 9, 81.0), ('Fido', 3, 27.0)],
             dtype=[('name', 'U10'), ('age', 'i4'), ('weight', 'f4')])
x['name']
x[['name', 'age']]

"""END"""
```

### The ix~()~ function

1.  The ix\_ function can be used to combine different vectors so as to
    obtain the result for each n-uplet.(类似于R中的 `expand.grid` 函数)
2.  For example, if you want to compute all the a+b\*c for all the
    triplets taken from each of the vectors a, b and c:

``` {python}
a = np.array([2,3,4,5])
b = np.array([8,5,4])
c = np.array([5,4,6,8,3])
ax,bx,cx = np.ix_(a,b,c)

result = ax+bx * cx
result

"""END"""
```

# 通用函数UNIVERSAL FUNCTIONS (UFUNC)

### 简介

1.  A universal function (or ufunc for short) is a function that
    operates on ndarrays in an element-by-element fashion.
2.  That is, a ufunc is a "vectorized" wrapper for a function that takes
    a fixed number of specific inputs and produces a fixed number of
    specific outputs.
3.  In NumPy, universal functions are instances of the numpy.ufunc
    class.
4.  Many of the built-in functions are implemented in compiled C code.

### 常见通用函数

1.  Math operations
2.  Trigonometric functions
3.  Floating functions
4.  。。。

# 统计功能

### Order statistics

  函数                                      功能
  ----------------------------------------- ---------------------------------------------------------------------------------------------
  amin(a\[, axis, out, keepdims\])          Return the minimum of an array or minimum along an axis.
  amax(a\[, axis, out, keepdims\])          Return the maximum of an array or maximum along an axis
  nanmin(a\[, axis, out, keepdims\])        Return minimum of an array or minimum along an axis, ignoring any NaNs.
  nanmax(a\[, axis, out, keepdims\])        Return the maximum of an array or maximum along an axis, ignoring any NaNs.
  ptp(a\[, axis, out\])                     Range of values (maximum - minimum) along an axis.
  percentile(a, q\[, axis, out, ...\])      Compute the qth percentile of the data along the specified axis
  nanpercentile(a, q\[, axis, out, ...\])   Compute the qth percentile of the data along the specified axis, while ignoring nan values.

### Averages and variances

  函数                                                   功能
  ------------------------------------------------------ ------------------------------------------------------------------------------
  median(a\[, axis, out, overwrite~input~, keepdims\])   Compute the median along the specified axis.
  average(a\[, axis, weights, returned\])                Compute the weighted average along the specified axis.
  mean(a\[, axis, dtype, out, keepdims\])                Compute the arithmetic mean along the specified axis.
  std(a\[, axis, dtype, out, ddof, keepdims\])           Compute the standard deviation along the specified axis.
  var(a\[, axis, dtype, out, ddof, keepdims\])           Compute the variance along the specified axis.
  nanmedian(a\[, axis, out, overwrite~input~, ...\])     Compute the median along the specified axis, while ignoring NaNs.
  nanmean(a\[, axis, dtype, out, keepdims\])             Compute the arithmetic mean along the specified axis, ignoring NaNs.
  nanstd(a\[, axis, dtype, out, ddof, keepdims\])        Compute the standard deviation along the specified axis,while ignoring NaNs.
  nanvar(a\[, axis, dtype, out, ddof, keepdims\])        Compute the variance along the specified axis, while ignoring NaNs.

注：几何平均数，调和平均数函数在scipy中

### Correlating

  函数                                               功能
  -------------------------------------------------- ---------------------------------------------------------
  corrcoef(x\[, y, rowvar, bias, ddof\])             Return Pearson product-moment correlation coefficients.
  correlate(a, v\[, mode\])                          Cross-correlation of two 1-dimensional sequences.
  cov(m\[, y, rowvar, bias, ddof, fweights, ...\])   Estimate a covariance matrix, given data and weights.

### Histograms

# Linear algebra (numpy.linalg)

### 常见矩阵运算- Matrix and vector products

\scriptsize

  函数                         功能
  ---------------------------- -------------------------------------------------
  dot(a, b)                    Dot product of two arrays.
  vdot(a, b)                   Return the dot product of two vectors.
  inner(a, b)                  Inner product of two arrays.
  outer(a, b)                  Compute the outer product of two vectors.
  linalg.matrix~power~(M, n)   Raise a square matrix to the (integer) power n.
  kron(a, b)                   Kronecker product of two arrays.

### 常见矩阵运算- Decompositions

\scriptsize

  函数                 功能
  -------------------- -------------------------------------------------------------------
  linalg.cholesky(a)   Cholesky decomposition.
  linalg.qr(a)         Compute the qr factorization of a matrix.
  linalg.svd(a)        Singular Value Decomposition.
  linalg.eig(a)        Compute the eigenvalues and right eigenvectors of a square array.
  scipy.linalg.lu(a)   Compute pivoted LU decompostion of a matrix.

### 常见矩阵运算- Norms and other numbers

\scriptsize

  函数                     功能
  ------------------------ ----------------------------------------------
  linalg.norm(x)           Matrix or vector norm.
  linalg.cond(x)           Compute the condition number of a matrix.
  linalg.det(a)            Compute the determinant of an array.
  linalg.matrix~rank~(M)   Return matrix rank of array using SVD method
  trace(a)                 Return the sum along diagonals of the array.

### 常见矩阵运算- Solving equations and inverting matrices

\scriptsize

  函数                       功能
  -------------------------- -----------------------------------------------------------------------
  linalg.solve(a, b)         Solve a linear matrix equation, or system of linear scalar equations.
  linalg.tensorsolve(a, b)   Solve the tensor equation a x = b for x.
  linalg.lstsq(a, b)         Return the least-squares solution to a linear matrix equation.
  linalg.inv(a)              Compute the (multiplicative) inverse of a matrix.
  linalg.pinv(a)             Compute the (Moore-Penrose) pseudo-inverse of a matrix.

# Random sampling (numpy.random)

### 简介

1. 下面这些函数主要用于随机抽样和生成随机数字，关于概率，分位点等计算见  `scipy.stats` 模块
2. 下面函数都以 `np.random.` 开始
3. 有很多功能相同名字不同的函数

### Simple random data

\scriptsize

  函数                                    功能
  --------------------------------------- -----------------------------------------------------------------------
  rand(d0, d1, ..., dn)                   Random values in a given shape.
  randn(d0, d1, ..., dn)                  Return a sample (or samples) from the "standard normal" distribution.
  randint(low\[, high, size, dtype\])     Return random integers from low (inclusive) to high (exclusive).
  random~integers~(low\[, high, size\])   Random integers of type np.int between low and high, inclusive.
  random~sample~(\[size\])                Return random floats in the half-open interval \[0.0, 1.0).
  random(\[size\])                        Return random floats in the half-open interval \[0.0, 1.0).
  ranf(\[size\])                          Return random floats in the half-open interval \[0.0, 1.0).
  sample(\[size\])                        Return random floats in the half-open interval \[0.0, 1.0).
  choice(a\[, size, replace, p\])         Generates a random sample from a given 1-D array
  bytes(length)                           Return random bytes.

### Permutations

\scriptsize

  函数             功能
  ---------------- ----------------------------------------------------------
  shuffle(x)       Modify a sequence in-place by shuffling its contents.
  permutation(x)   Randomly permute a sequence, or return a permuted range.

### Distribution

\scriptsize

  函数                                             功能
  ------------------------------------------------ -----------------------------------------------------------------------------------------------------------------------
  beta(a, b\[, size\])                             Draw samples from a Beta distribution.
  binomial(n, p\[, size\])                         Draw samples from a binomial distribution.
  chisquare(df\[, size\])                          Draw samples from a chi-square distribution.
  dirichlet(alpha\[, size\])                       Draw samples from the Dirichlet distribution.
  exponential(\[scale, size\])                     Draw samples from an exponential distribution.
  f(dfnum, dfden\[, size\])                        Draw samples from an F distribution.
  gamma(shape\[, scale, size\])                    Draw samples from a Gamma distribution.
  geometric(p\[, size\])                           Draw samples from the geometric distribution.
  gumbel(\[loc, scale, size\])                     Draw samples from a Gumbel distribution.
  hypergeometric(ngood, nbad, nsample\[, size\])   Draw samples from a Hypergeometric distribution.
  laplace(\[loc, scale, size\])                    Draw samples from the Laplace or double exponential distribution with specified location (or mean) and scale (decay).
  logistic(\[loc, scale, size\])                   Draw samples from a logistic distribution.
  lognormal(\[mean, sigma, size\])                 Draw samples from a log-normal distribution.
  logseries(p\[, size\])                           Draw samples from a logarithmic series distribution.

### Distributions

\scriptsize

  函数                                             功能
  ------------------------------------------------ -----------------------------------------------------------------------------------
  multivariate~normal~(mean, cov\[, size, ...\])   Draw random samples from a multivariate normal distribution.
  negative~binomial~(n, p\[, size\])               Draw samples from a negative binomial distribution.
  noncentral~chisquare~(df, nonc\[, size\])        Draw samples from a noncentral chi-square distribution.
  noncentral~f~(dfnum, dfden, nonc\[, size\])      Draw samples from the noncentral F distribution.
  normal(\[loc, scale, size\])                     Draw random samples from a normal (Gaussian) distribution.
  pareto(a\[, size\])                              Draw samples from a Pareto II or Lomax distribution with specified shape.
  poisson(\[lam, size\])                           Draw samples from a Poisson distribution.
  power(a\[, size\])                               Draws samples in \[0, 1\] from a power distribution with positive exponent a - 1.
  rayleigh(\[scale, size\])                        Draw samples from a Rayleigh distribution.
  standard~cauchy~(\[size\])                       Draw samples from a standard Cauchy distribution with mode = 0.
  standard~exponential~(\[size\])                  Draw samples from the standard exponential distribution.
  standard~gamma~(shape\[, size\])                 Draw samples from a standard Gamma distribution.
  standard~normal~(\[size\])                       Draw samples from a standard Normal distribution (mean=0, stdev=1).
  standard~t~(df\[, size\])                        Draw samples from a standard Student's t distribution with df degrees of freedom.
  triangular(left, mode, right\[, size\])          Draw samples from the triangular distribution over the interval \[left, right\].
  uniform(\[low, high, size\])                     Draw samples from a uniform distribution.
  vonmises(mu, kappa\[, size\])                    Draw samples from a von Mises distribution.
  wald(mean, scale\[, size\])                      Draw samples from a Wald, or inverse Gaussian, distribution.
  weibull(a\[, size\])                             Draw samples from a Weibull distribution.
  zipf(a\[, size\])                                Draw samples from a Zipf distribution.

### Random generator

\scriptsize

  函数                    功能
  ----------------------- --------------------------------------------------------------------
  RandomState(\[seed\])   Container for the Mersenne Twister pseudo-random number generator.
  seed(\[seed\])          Seed the generator.
  get~state~()            Return a tuple representing the internal state of the generator.
  set~state~(state)       Set the internal state of the generator from a tuple.

# Functions and Methods Overview

### Functions and Methods Overview

1.  **Array Creation**:

    arange, array, copy, empty, empty~like~, eye, fromfile,
    fromfunction,identity, linspace, logspace, mgrid, ogrid, ones,
    ones~like~, r, zeros, zeros~like~

2.  **Conversions**:

    ndarray.astype, atleast~1d~, atleast~2d~, atleast~3d~, mat

3.  **Manipulations**:

    array~split~, column~stack~, concatenate, diagonal, dsplit, dstack,
    hsplit,hstack, ndarray.item, newaxis, ravel, repeat, reshape,
    resize, squeeze, swapaxes, take, transpose, vsplit, vstack

### Functions and Methods Overview

1.  **Questions**:

    all, any, nonzero, where

2.  **Ordering**:

    argmax, argmin, argsort, max, min, ptp, searchsorted, sort

3.  **Operations**:

    choose, compress, cumprod, cumsum, inner, ndarray.fill, imag, prod,
    put, putmask,real, sum

4.  **Basic Statistics**:

    cov, mean, std, var

5.  **Basic Linear Algebra**:

    cross, dot, outer, linalg.svd, vdot

# Other Subpackages

### Numpy中的其他常用模块

1.  String operations
2.  Datetime Support Functions
3.  Discrete Fourier Transform (numpy.fft)
4.  Financial functions
5.  Functional programming
6.  Logic functions
7.  Mathematical functions
8.  Matrix library (numpy.matlib)
9.  numpy.polynomial package
10. 具体用法和更多模块可以参考 Numpy reference\[numpy-ref.pdf\]中的
    Routines 内容


<!-- # 参考文献 -->
[//]: # (\bibliography{Bibfile})
