---
title: "statsmodels 用法"
author: "Jin"
date: "2020年6月"
institute: 中南财经政法大学统计与数学学院
csl: ./style/chinese-gb7714-2015-numeric.csl
css: ./style/markdown.css
bibliography: [./Bibfile.bib]
eqnPrefixTemplate: ($$i$$)
link-citations: true
linkReferences: true
chapters: true
tableEqns: false
autoEqnLabels: false
classoption: "aspectratio=1610"
---

```{r setup, echo=F, purl=F}
knitr::opts_knit$set(root.dir = getwd())
knitr::opts_chunk$set(echo = TRUE, results = 'hide')
knitr::opts_chunk$set(warning = FALSE, message=FALSE)
knitr::opts_chunk$set(fig.align="center"
                      ## ,out.width="0.9\\textwidth" # latex
                      ,out.width="80%" # for both latex and html
                      ,fig.width=5, fig.height=3
                      )
```

```{r prepare, echo=F, purl=F}
rm(list=ls())
options(digits=4)
options(scipen=100)
graphics.off()
Sys.setlocale("LC_ALL", "Chinese")
library(reticulate)
```



# Introduction

### Facts

1.  Initial release: 2009
2.  Website: <https://www.statsmodels.org/stable/index.html>
3.  History:

```{=html}
<!-- -->
```
1.  The models module of scipy.stats was originally written by Jonathan
    Taylor.
2.  For some time it was part of scipy but was later removed.
3.  During the Google Summer of Code 2009, statsmodels was corrected,
    tested, improved and released as a new package.
4.  Since then, the statsmodels development team has continued to add
    new models, plotting tools, and statistical methods.

### What is statsmodels

1.  statsmodels is a Python module that provides classes and functions
    for the estimation of many different statistical models,
2.  as well as for conducting statistical tests, and statistical data
    exploration.
3.  An extensive list of result statistics are available for each
    estimator.
4.  The results are tested against existing statistical packages to
    ensure that they are correct.

### A simple example using ordinary least squares

``` {.python}
import numpy as np
import statsmodels.api as sm
import statsmodels.formula.api as smf

dat = sm.datasets.get_rdataset("Guerry", site="C:/github_repo/Rdatasets", package="HistData").data
results = smf.ols('Lottery ~ Literacy + np.log(Pop1831)', data=dat).fit()
print(results.summary())

nobs = 100
X = np.random.random((nobs, 2))
X = sm.add_constant(X)
beta = [1, .1, .5]
e = np.random.random(nobs)
y = np.dot(X, beta) + e

results = sm.OLS(y, X).fit()
print(results.summary())
```

# 使用流程

## 导入路径和结构

### 两种导入方式

1.  从statsmodel导入函数和类的两种方法:
    1.  用于交互使用的API导入：允许选项卡完成
    2.  程序的直接导入：避免导入不必要的模块和命令

### 用于交互使用的API导入

1.  对于交互式使用，建议导入为：
    `import statsmodels.api as sm`
2.  导入statsmodels.api将加载大部分statsmodel的公共部分。
3.  这使得大多数函数和类可以在一到两层中方便地使用，而不会使“sm”名称空间过于拥
    挤。
4.  用 `dir(sm)` 列出函数。

### 程序的直接导入

1.  statsmodels子模块按主题排列(例如离散选择模型的离散，或者时间序列分析的tsa)。

函数和类：

`from statsmodels.regression.linear_model import OLS, WLS`
`from statsmodels.tools.tools import rank, add_constant`

模块

`from statsmodels.datasets import macrodata`
`import statsmodels.stats import diagnostic`

模块的别名

`import statsmodels.regression.linear_model as lm`
`import statsmodels.stats.diagnostic as smsdia`
`import statsmodels.stats.outliers_influence as oi`

## 使用R样式公式拟合模型

### R样式公式和 `pasty`

1.  statsmodels允许用户使用R样式公式拟合统计模型。
2.  在内部，statsmodels使用patsy包将公式和数据转换为模型拟合中使用的矩阵。
3.  patsy是一个Python软件包，用于描述统计模型（尤其是线性模型或具有线性成分的模
    型）并构建设计矩阵。 它受到R和S中使用的微型语言公式的紧密启发并与之兼容。
4.  例如，如果我们有一些变量y，并且想要将其与其他变量x，a，b以及a和b的相互作用进
    行回归，则只需编写：
    `patsy.dmatrices("y ~ x + a + b + a:b", data)`

### `dmatrices`

1.  将类别Region变量拆分为一组指标变量。
2.  在外生回归矩阵中添加一个常数。
3.  返回的是Pandas DataFrames而不是简单的numpy数组。这很有用，因为在报告结果
    时，DataFrames允许statsmodels继承元数据（例如，变量名）。

### 模型拟合和总结

1.  在statsmodels中拟合模型通常涉及3个简单步骤：
    1.  使用模型类来描述模型。
    2.  使用类方法拟合模型。
    3.  使用汇总方法检查结果。
    4.  输入 `dir(res)` 以获取完整的属性列表。

``` {.python}
mod = sm.OLS(y, X)    # Describe model
res = mod.fit()       # Fit model
print(res.summary())   # Summarize model
```

# 主要 statsmodels API

### main statsmodels API

1.  `statsmodels.api`: Cross-sectional models and methods. Canonically
    imported using `import statsmodels.api as sm.`
2.  `statsmodels.tsa.api`: Time-series models and methods. Canonically
    imported using `import statsmodels.tsa.api as tsa.`
3.  `statsmodels.formula.api`: A convenience interface for specifying
    models using formula strings and DataFrames. This API directly
    exposes the from~formula~ class method of models that support the
    formula API. Canonically imported using
    `import statsmodels.formula.api as smf`
4.  The API focuses on models and the most frequently used statistical
    test, and tools.

### `statsmodels.api`

1.  Regression
2.  Imputation
3.  Generalized Estimating Equations
4.  Generalized Linear Models
5.  Discrete and Count Models
6.  Multivariate Models
7.  Misc Models
8.  Graphics
9.  Tools

### `statsmodels.tsa.api`

1.  Statistics and Tests
2.  Univariate Time-Series Analysis
3.  Exponential Smoothing
4.  Multivariate Time Series Models
5.  Filters and Decompositions
6.  Markov Regime Switching Models
7.  Time-Series Tools
8.  X12/X13 Interface

### `statsmodels.formula.api`

1.  The function descriptions of the methods exposed in the formula API
    are generic. See the documentation for the parent model for details.
    -   gls(formula, data\[, subset, drop~cols~\])
    -   wls(formula, data\[, subset, drop~cols~\])
    -   ols(formula, data\[, subset, drop~cols~\])
    -   mixedlm(formula, data\[, re~formula~, ...\])
    -   glm(formula, data\[, subset, drop~cols~\])
    -   mnlogit(formula, data\[, subset, drop~cols~\])
    -   logit(formula, data\[, subset, drop~cols~\])
    -   probit(formula, data\[, subset, drop~cols~\])
    -   poisson(formula, data\[, subset, drop~cols~\])
    -   negativebinomial(formula, data\[, subset, ...\])
    -   quantreg(formula, data\[, subset, drop~cols~\])
    -   ordinal~gee~(formula, groups, data\[, subset, ...\])
    -   nominal~gee~(formula, groups, data\[, subset, ...\])
    -   gee(formula, groups, data\[, subset, time, ...\])
    -   glmgam(formula, data\[, subset, drop~cols~\])


# 数据集包

### 数据集包

1.  statsmodels提供用于示例、教程、模型测试等的数据集（即数据和元数据）
2.  加载statsmodels中的可用数据集
3.  使用Stata中的数据集
4.  使用R中的数据集

### statsmodels可用数据集

1.  可用数据集列表：
    `Anaconda3\Lib\site-packages\statsmodels\datasets`
2.  加载数据集：`sm.datasets.datasets_name.load_pandas()`
3.  将数据加载为pandas对象：`load_pandas()` 方法
4.  数据集对象遵循束模式，完整的数据集在“data”属性中可用。
5.  大多数数据集都在属性“endog”和“exog”中保存了方便的数据表示形式。
6.  但是，单变量数据集没有“exog”属性。
7.  变量名可以通过键入以下内容获得：`endog_name`和`exog_name`

### 例子

``` {.python}
import statsmodels.api as sm
dat = sm.datasets.longley.load_pandas()

dat.data
dat.endog
dat.exog
dat.endog_name
dat.exog_name
dat.names
```

### 使用Stata中的数据集

1.  `webuse(data[, baseurl, as_df])` : Download and return an example
    dataset from Stata.

``` {.python}
import statsmodels.api as sm

auto=sm.datasets.webuse('auto')
```

### 使用R中的数据集

1.  The Github
    [Rdatasets](https://github.com/vincentarelbundock/Rdatasets) 项目可以访问R的核心数据包和许多其他常见R包中的可用数据集。
2.  使用`get_rdataset`函数，所有这些数据集都可用于statsmodels。
3.  实际数据可以通过“data”属性访问。
4.  `__doc__` 属性可以查看数据帮助信息。
5.  由于 Github 网站文件库被封，可以采取以下方法解决：
    1.  把 [Rdatasets](https://github.com/vincentarelbundock/Rdatasets)项目下载到本地计算机。
    2.  修改 `Anaconda3\Lib\site-packages\statsmodels\datasets\utils.py`中的 `get_rdataset` 函数。

``` {.python}
import statsmodels.api as sm

iris=sm.datasets.get_rdataset('iris', site="C:/github_repo/Rdatasets")
iris.data
print(iris.__doc__)
```


<!-- # 参考文献 -->
[//]: # (\bibliography{Bibfile})